Update Meeting: 03-11-2021

### 2021-11-22 analysis/update_meeting_03_11_2021.Rmd
wflow_publish("analysis/update_meeting_03_11_2021.Rmd")

Decided to run MVMR as follows: $Y_p \sim X_i + Y_i + X_p$, using instruments for ${X_i, Y_i, X_p}$. In this MR, the coefficient for $X_i$ would represent the *direct* $X_i \rightarrow Y_p$ causal effect and hopefully this would be close to zero in most cases. Running simply the $Y_p \sim X_i$ MR (with only $X_i$ instruments), would give you the $X_i \rightarrow Y_p$ *total* effect. Then we can compare the direct and total effect results.

A summary of the MVMR results are below, filtered to only traits with abs(correlation) < 0.8, and then BF significant with $X_i$.

@@

num_tests <- dim(household_MVMR_summary.DT)[1]
household_MVMR_BF.DT <- household_MVMR_summary.DT %>% filter(xi_meta_pval < 0.05/num_tests) %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, yi_meta_beta, yi_meta_pval, xi_meta_beta, xi_meta_pval, xp_meta_beta, xp_meta_pval) %>% unique()

@@

@@

plot_data <-  inner_join(household_MR_summary_sub, household_MVMR_summary.DT)  %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, IVW_meta_beta, IVW_meta_se, yi_meta_beta, yi_meta_se, xi_meta_beta, xi_meta_se, xp_meta_beta, xp_meta_se) %>% unique()

@@

>>>>

Decided to run MVMR as follows: $Y_p \sim X_i + Y_i + X_p$, using instruments for ${X_i, Y_i, X_p}$. In this MR, the coefficient for $X_i$ would represent the **direct** $X_i \rightarrow Y_p$ causal effect and hopefully this would be close to zero in most cases. Running simply the $Y_p \sim X_i$ MR (with only $X_i$ instruments), would give you the $X_i \rightarrow Y_p$ **total** effect. Then we can compare the direct and total effect results.

A summary of the MVMR results are below, filtered to only traits with `abs(correlation) < 0.8`, and then BF significant with $X_i$.

@@

num_tests <- dim(household_MVMR_summary.DT)[1]
household_MVMR_BF.DT <- household_MVMR_summary.DT %>% filter(xi_meta_pval < 0.05/num_tests*2) %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, yi_meta_beta, yi_meta_pval, xi_meta_beta, xi_meta_pval, xp_meta_beta, xp_meta_pval) %>% unique()

@@

The plot below displays the direct effect (from MVMR) vs indirect effects (from univariate MR) of $X_i$ on $Y_p$ for all combinations of traits with `abs(correlation) < 0.8` (i.e. the data above plus non-significant results). Total number of tests is `r dim(household_MVMR_summary.DT)[1]/2`. The second plot shows only BF-significant results according to the MVMR.

@@

  plot_data <-  inner_join(household_MR_summary_sub, household_MVMR_summary.DT)  %>% dplyr::select(exposure_ID, outcome_ID, exposure_description, outcome_description, corr_traits, IVW_meta_beta, IVW_meta_se, yi_meta_beta, yi_meta_se, xi_meta_beta, xi_meta_se, xi_meta_pval, xp_meta_beta, xp_meta_se) %>% unique()

sig_plot_data <- plot_data %>% filter(xi_meta_pval < 0.05/num_tests*2)

@@

plot_sig <-
  ggplot2::ggplot(data = sig_plot_data, ggplot2::aes(x = IVW_meta_beta,
                                                 y = xi_meta_beta)) +
  geom_point(alpha = 3 / 4) +
  geom_smooth(
    mapping = aes(x = IVW_meta_beta, y = xi_meta_beta),
    method = "lm",
    se = FALSE,
    formula = y ~ x + 0,
    fullrange = TRUE,
    color = custom_col[4]
  ) +
  ggplot2::geom_errorbar(
    ggplot2::aes(ymin = xi_meta_beta - xi_meta_se, ymax = xi_meta_beta + xi_meta_se),
    colour = "grey",
    width = 0
  ) +
  ggplot2::geom_errorbarh(
    ggplot2::aes(xmin = IVW_meta_beta - IVW_meta_se, xmax = IVW_meta_beta + IVW_meta_se),
    colour = "grey",
    height = 0
  ) +
  theme_minimal() +
  ggplot2::labs(
    x = expression(Indirect~effect~of~X[i]~on~Y[p]~(from~univariate~MR)),
    y = expression(Direct~effect~of~X[i]~on~Y[p]~(from~MVMR))
  ) + 
  geom_abline(slope=1, intercept=0)

plot_sig
```
